---
title: "ESM204_Assignment3"
author: "Bridget Gibbons, Chase Brewster"
date: "5/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = FALSE}
#Load packages

library(tidyverse)
```

```{r load data, include = FALSE}
#read in the data

survey_data <- read_csv("HW3Data.csv")
```

###1.

Data colums are: 1) risk - level of risk reduction, 2) bid - annual household payment, 3) vote - 1 is yes, 0 is no, 4) NEP - measurement of environmental concern, 5) income, 6) age

```{r part 1, include = FALSE}
#run a linear regression

whales_lm <- lm(vote ~ risk + bid + NEP + income + age, data = survey_data)
summary(whales_lm)

#save all the coefficients

intercept <- whales_lm$coefficients[1]
risk <- whales_lm$coefficients[2]
bid <- abs(whales_lm$coefficients[3])
nep <- whales_lm$coefficients[4]
inc_onepercent <- whales_lm$coefficients[5]
inc_poor <- whales_lm$coefficients[6]
inc_rich <- whales_lm$coefficients[7]
inc_veryrich <- whales_lm$coefficients[8]
age_50 <- whales_lm$coefficients[9]
age_40 <- abs(whales_lm$coefficients[10])
age_60 <- abs(whales_lm$coefficients[11])
age_30 <- whales_lm$coefficients[12]
```

**Regression Model Output:**  
  
Probability(Voting Yes) = `r round(intercept, digits = 4)` + `r round(age_30, digits=4)`(Age to 30) - `r round(age_40, digits=4)`(Age to 40) + `r round(age_50, digits=4)`(Age to 50) - `r round(age_60, digits=4)`(Age to 60) + `r round(inc_onepercent, digits=4)`(Income One Percent) + `r round(inc_poor, digits=4)`(Income Poor) + `r round(inc_rich, digits=4)`(Income Rich) + `r round(inc_veryrich,digits=4)`(Income Very Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + $7.4x10^{-4}$(Risk Reduction)

**Explanation of Coefficients:**

Age:  

*To 30:* All else equal, the average person in the Age to 30 bracket is expected to have a 0.0204 higher probability of voting yes on the VSR program than the average person in the Age over Sixty bracket.  

*To 40:* All else equal, the average person in the Age to 40 bracket is expected to have a 0.0201 lower probability of voting yes on the VSR program than the average person in the Age over Sixty bracket.  

*To 50:* All else equal, the average person in the Age to 50 bracket is expected to have a 0.01 higher probability of voting yes on the VSR program than the average person in the Age over Sixty bracket.  

*To 60:* All else equal, the average person in the Age to 60 bracket is expected to have a 0.0162 lower probability of voting yes on the VSR program than the average person in the Age over Sixty bracket.

Income:   

*Poor:* All else equal, the average person in the Poor bracket is expected to have a 0.0027 higher probability of voting yes on the VSR program than the average person in the Middle bracket.  

*Rich:* All else equal, the average person in the Rich bracket is expected to have a 0.0075 higher probability of voting yes on the VSR program than the average person in the Middle bracket.  

*Very Rich:* All else equal, the average person in the Very Rich bracket is expected to have a 0.0468 higher probability of voting yes on the VSR program than the average person in the Middle bracket.  

*One Percent:* All else equal, the average person in the One Percent bracket is expected to have a 0.0088 higher probability of voting yes on the VSR program than the average person in the Middle bracket.

NEP: All else equal, every 1 unit increase in environmental concern is expected to increase the average person’s probability of voting yes on the VSR program by 0.0159.

Bid: All else equal, every $1 increase in the program cost is expected to reduce the average person’s probability of voting yes on the VSR program by 0.0011.

Risk: All else equal, every 1% increase in whale collision risk reduction is expected to increase the average person’s probability of voting yes on the VSR program by $7×10^{−4}$.


Reference levels in the model are the middle income range and the age of over_sixty.

###2. 
One prevented whale death equates to a 4% risk reduction. In order to calculate the value of a single prevented whale death, the above identified linear model was used to calculate the bid (willingness to pay) for a 0% reduction in risk and a 4% reduction in risk. The difference in the bids represents the value of one prevented whale death. The model incorporates p(yes) = 0.5, which will give us the median willingness to pay, mode age (to_thirty), mode income (one_percent), and the average NEP (38.366).

WTP:
$0.5 = `r round(intercept, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_onepercent, digits=4)`(Income~One~Percnet) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(0)$

```{r, include = FALSE}
#A reduction by 4% saves a single whale, according to the ratio in the survey. Find the WTP for vessel speed reduction programs of 0 (risk) and compare to the WTP for vessel speed reduction of 4 (risk) to find WTP for saving one whale.

#p(yes) = 0.5 for median WTP:
med_vote <- 0.5
#0.5

#average of the NEP:
nep_avg <- mean(survey_data$NEP)
#38.366

#bid if risk reduction is zero:
bid_zero <- (intercept + age_30 + inc_onepercent + (nep*nep_avg) + (risk*0) - 0.5)/bid
bid_zero
#240.77

#Bid if risk reduction is four:
bid_four <- (intercept + age_30 + inc_onepercent + (nep*nep_avg) + (risk*4) - 0.5)/bid
bid_four
#243.55
```

Value of one whale = 243.55 - 240.77 = $2.78 per Santa Barbara resident

###3.

Three arbitrary individuals were selected from the dataset. The linear model was rearranged to output the bid, or willingness to pay, and the model inputs for age, income, and NEP were changed to match the respondent demographics. Risk reduction was set to 60% for each individual. 

```{r, include = FALSE}
#Pick three random respondents from the table: 47,135,371

risk60_40 <- slice(survey_data, 40)

risk60_135 <- slice(survey_data, 135)

risk60_372 <- slice(survey_data, 372)
```

Individual 40:
$`r med_vote` = `r round(intercept, digits = 4)` + `r round(age_50, digits=4)`(Age~to~50) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

Individual 135:
$`r med_vote` = `r round(intercept, digits = 4)` + `r round(inc_onepercent, digits=4)`(Income~One~Percent) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

Individual 372:
$`r med_vote` = `r round(intercept, digits = 4)` + `r round(age_50, digits=4)`(Age~to~50) + `r round(inc_veryrich, digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$


```{r, include = FALSE}

bid_40 <- (intercept + age_50 + inc_poor + (nep*34) + (risk*60) - med_vote)/bid
bid_40
#202.32

bid_135 <- (intercept + inc_onepercent + (nep*50) + (risk*60) - med_vote)/bid
bid_135
#435.92

bid_372 <- (intercept + inc_veryrich + age_50 + (nep*48) + (risk*60) - med_vote)/bid
bid_372
#451.08

```
Solving for bid:

**Individual 40:** 202.32  

**Individual 135:** 435.92  

**Individual 372:** 451.08

###4
The mean WTP (bid) for a VSR program among all Santa Barbara households was calculated using a probability of voting yes of 0.5, the mode age (to_thirty), the mode income (one_percent), and the average NEP (38.366) in the linear model.

$`r med_vote` = `r round(intercept, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(age_30, digits=4)`(Age~to~30) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$  

```{r, include = FALSE}

avg_bid_sb <- (intercept + age_30 + inc_onepercent + (nep*nep_avg) + (risk*60) - med_vote)/bid
avg_bid_sb

#282.52

```

Median WTP in Santa Barbara households is $282.52

###5
The total benefit for a VSR program is equal to the number of households (150,000) times the mean WTP calculated above. 

Total Benefits:
$282.52*150,000 = 42,378,000 $

Cost = $7 million

Net Benefits:
$42,378,000 - $7,000,000 = $35,378,000

###6
The total benefit is greater than the $7,000,000 implementation costs, so the VSR program should be implemented.

##7
Z* is equal to the costs ($1,000) divided by the tons of carbon (20) and is equal to $50 per ton.

20xZ* = $1,000
Z* = $50/ton

###8

a) If a risk reduction of 4% saves 1 whale, 15 whales would be saved by the program

b) $42,378,000 because the net social benefit is equal to the benefits of the whales to santa barbara plus the carbon reduction benefits minus the implementation costs. At Z*, the carbon reduction benefits are equal to the program implementation costs, so those cancel out and we are left with the total benefit calculated in Part 5.

